---
title: "Pollen competition"
author: "Fernando Cagua, Hugo Marrero, Jason Tylianakis, Daniel Stouffer"
date: "19 December 2017"
output: pdf_document
bibliography: references.bib
csl: ecology-letters.csl
header-includes:
- \usepackage{booktabs}
- \usepackage{tabularx}
- \usepackage{setspace}
---

\doublespacing

```{r, echo = F, message = F}
library(magrittr)
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE, 
  warning = FALSE
)
options("xtable.include.rownames" = FALSE,
        "xtable.comment" = FALSE,
        "xtable.size" = "small")
```

## Abstract

## Introduction

## Methods

### Data collection

We performed sampling in three locations, each located in a distinct phytogeograohic region of the Argentinean Pampas (Figure 1). 
The sampling locations are distributed across a gradient of precipitation with a humid temperate climate in the eastward Flooding Pampas and an drier as we move inland and westward towards the Pampean Grasslands.
Across all sites, annual mean temperatures range between 14.8 and 15.8 Celsius. 

In each location, we sampled two restored and two agricultural fragments. 
However, due to the lack of available sites in the Flooding Pampas, it was only possible to sample one restored site.
We visited each fragment monthly between November and March between 2010 and 2011. 
In each visit we gathered information related to abundance, plant-pollinator visitation, pollen transfer, and pollen deposition.
For more information about the study sites and data collection protocols can be found in previous publications [@Marrero2014; @Marrero2016; @Marrero2017].

```{r, results='asis'}
readr::read_csv("variables_sampled.csv") %>%
  dplyr::rename(` ` = Var) %>%
  xtable::xtable(caption = "Variables sampled", align = c("l", "l", "X")) %>%
  print(booktabs = TRUE, tabular.environment = "tabularx", width = "\\linewidth")
```

### Data analysis

## Results

## Discussion

## References