---
title: "The pollination trade-off"
author: "Fernando Cagua^1^, Hugo Marrero^2^, Jason Tylianakis^1^, Daniel Stouffer^1^"
bibliography: phd-bibliography/pollen-competition.bib
csl: ecology-letters.csl
output: 
  bookdown::pdf_document2: 
    keep_tex: yes
    number_sections: false
toc: false
header-includes:
  - \usepackage{booktabs}
  - \usepackage{setspace}
  - \usepackage{lineno}
  - \usepackage{xr}
  - \newcommand{\R}[1]{\label{#1}\linelabel{#1}}
  - \newcommand{\lr}[1]{page~\pageref{#1}, line~\lineref{#1}}
  - \externaldocument[S-]{supp-info}
---

\linenumbers
\doublespacing
 
^1^ Centre for Integrative Ecology, School of Biological Sciences, University of Canterbury, Private Bag 4800, Christchurch 8041, New Zealand

^2^ Centro de Recursos Naturales Renovables de las Zonas Semiáridas, CONICET, Camino de la Carrindanga Km. 7, 8000 Bahía Blanca, Argentina

**Author for correspondence:** E. Fernando Cagua (efc29@uclive.ac.nz) - Centre for Integrative Ecology, School of Biological Sciences, University of Canterbury, Private Bag 4800, Christchurch 8140, New Zealand

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, 
                      warning = FALSE)
```

```{r draft-details}
drake::loadd(msc_wordcount, abs_wordcount)
```

**Manuscript length:** `r format(msc_wordcount$n_char_tot_stri, big.mark = ",")` characters; approx. `r format(round(msc_wordcount$n_words_stri/10)*10-90, big.mark = ",")` words (includes draft info, excludes front-matter, abstract, awknoledgements, references, and figure/table legends); `r msc_wordcount$reading_time_stri` reading time.

**Abstract length:** `r abs_wordcount$n_words_stri` words.

\clearpage

# Draft Info

## Changes

* Now using AICc instead of AIC to evaluate models (no change in results)
* Traits: (*a*) to keep signature of phenology, absolute flower counts (log-transformed) are used instead of relative abundances. Also, now (*b*) columns in the trait matrix for phenology are scaled so that it doesn't count 5 times as much as other traits.
* The absolute/relative/gain terminology was confusing. Now  I'm modelling open and bagged pollen counts directly, using the bagged as a sort of control.
* Introduced visit efficacy instead of the share in the pollen pool as a predictor in the model. Turns out to be a better predictor and also incorporates quantitative visitation data which was missing from the previous models. 
* Now I use a PCA of the explanatory variables to understand different plant strategies and give a deeper teratment to the facilitation/competition question. 

## To do

* Spell-check everything & check that reference list is correct.
* Choose journal (the manuscript is written for an interdisciplinary journal). 

## Questions

Feel free to come back here after you've read the manuscript. Thanks!

* Do you know of a reasonable reference for the claims in \lr{ref-needed-possitive-loops}?
* I got feedback that "denote" in \lr{denote} might be overly formal? 
* Hugo. Some journals require publication of data alongside with the paper. Is it OK to consider those? Or you preffer to keep data private?

\clearpage

# Abstract

`r paste(drake::readd(abstract), collapse = "")`

**Keywords:** X, Y, Z

\clearpage

# Introduction

Animal pollination systems play a disproportionally important role in food production and maintenance of global biodiversity [@bascompte_plant-animal_2007; @klein_importance_2007; @ollerton_how_2011]. 
Alongside the direct mutualisms between plants and their pollinators, \R{ref-needed-possitive-loops} the support to biodiversity is substantially achieved as a result of inter-species facilitation—the positive feedback loops that exist between plants that share pollinators or pollinators that share plant. 
Facilitation is able to promote species coexistence because it offsets the effects of direct competition for resources [@stachowicz_mutualism_2001]. 
Indeed, studies that are fundamental to our current knowledge of mutualism predict that the maximum number of coexisting species---an upper limit to biodiversity---is achieved when the number of mutualistic partners is maximised in a community [@bastolla_architecture_2009]. 
And yet, not every pollinator species in natural communities interacts with every species of plant. 

One of the possible explanations is that sharing a mutualistic partner also has costs [@waser_interspecific_1978]. 
As soon as there are costs, plants start competing for mutualistic partners, even if this competition is not strong enough to produce a **net** negative effect in fitness.
The possible detrimental effects of competition for pollinators have been long recognised [@robertson_philosophy_1895] and widely documented both experimentally [@lewis_experimental_1961] and theoretically [@levin_competition_1970]. 
Multiple factors (like the number of shared pollinators, the plants' relative abundance, visitor fidelity, and traits) have been shown to modulate the strength of competition and, in turn, the fitness outcome [@bobisud_pollinator_1975; @mitchell_new_2009; @morales_interspecific_2008; @feinsinger_effects_1987; @campbell_pollinator_1985]. 
It is currently clear that the effects of these factors on fitness have the potential to drive the evolution of flower phenologies, morphologies and reproductive strategies, as well as ecological differentiation [@mitchell_new_2009; @caruso_competition_2000]. 

The main two mechanisms through which these ecological factors can affect plant fertilization are: (*i*) by reducing the number of conspecific pollen grains on stigmas or (*ii*) by increasing the deposition of heterospecific pollen [@morales_interspecific_2008]. 
For example, plants with more attractive flowers might reduce the number of visits---and hence the amount of conspecific pollen deposited by animals---for those less attractive [@yang_pollinator_2011]. 
We denote \R{denote} this as the *quantity* of the pollination service. 
More subtly, when plants share pollination partners, even receiving a visit might not translate into fertilization due to interspecific pollen transfer [@campbell_mechanism_1985]. 
Interspecific pollen transfer affects the pollination service because a focal plant might receive pollen from a different species, or conversely, pollen from the focal plant might be lost to different species. 
Generally speaking, the higher the proportion of conspecific relative to heterospecific pollen the higher the *quality* of the pollination service. 
With few exceptions [@rathcke_interactions_1988; @lopezaraizamikel_impact_2007; @tur_evaluating_2016; @hegland_relative_2009; @aizen_reproductive_2010], most of what we know about the quantity and quality of pollination---and its relationship to competition vs. facilitation in pollination systems---is based mostly on studies of two plant species. 
However, competitive interactions between species pairs do not occur in isolation. 

Instead, co-flowering plants often occur in communities in which multiple intertwined factors operate simultaneously and lead to emergent phenomena not observed at smaller scales. 
For instance, recent empirical evidence suggests that plants with flowering traits that are original relative to others in the community have generally fewer interaction partners [@coux_linking_2016]. 
This evidence is aligned with the notion that a species that interacts with few species does so strongly with each of them; in contrast, a species that interacts with a large number of species does so comparatively weakly. 
If evolutionary specialisation occurs by changing traits to focus on fewer but better partners [@caruso_competition_2000], we should expect a reduction of competition for pollination in species with original traits and an increase in species with a large number of interaction partners [@gibson_can_2012]. 
Alternatively, it might also be the case that flower/pollen density is the dominant force driving pollen transfer [@seifan_outcome_2014]. 
Abundant plant species might experience a dilution of available pollinators [@feinsinger_effects_1987] but might also receive more effective visits by capitalising on a larger share of both the pollen being transported by pollinators and their visits compared to other species. 
In this case, a potential reduction in the quantity of pollination (the amount of conspecific pollen received) could be compensated by an increase in the quality of pollination (the amount of conspecific relative to heterospecific pollen received). 
To complicate things more, autonomous self-pollination can also influence a plant sensitivity to competition for pollination because it can ensure reproduction even in the absence of outcross pollen [@kalisz_benefits_2003]. 
Altogether, multiple, potentially correlated, factors may act simultaneously.
Therefore, understanding the relative importance of these factors and how plant species harness them to minimise competition for pollination within their communities requires extensive data. 
Both describing the structure of the interactions and the pollen flows that occur within it. 

Here, we explore competition for pollination at the community scale using comprehensive empirical data from eleven plant-pollinator communities in the Argentinian Pampas. 
First, we ask the extent to which the outcome of animal-mediated pollination is facilitation or competition. 
If facilitation is the primary outcome we would expect (*i*) a larger amount of conspecific pollen deposited in stigmas when flowers are open to animal pollinators than that due to self-pollination and (*ii*) an overall positive relationship between the amount of conspecific pollen deposited in stigmas and the heterospecific pollen [@tur_evaluating_2016]. 
Second, we investigate the relative contribution that four ecological factors, that together, describe the role of the plant species and the ecological context, have on competition for pollination. 
We hypothesize that competition for pollination at the community level should increase as the number of pollination partners a plant has increases (species degree).
However, other factors like the plant's functional originality, its relative abundance, and the visit effectiveness, should have the potential to compensate this increase in competition by improving either the quality or the quality of pollination. 
Third, and finally, we explore how plants differ in how they use these factors to minimise competition. 
Although some plant species might be widespread across multiple locations, competition for pollinators occurs at the community level and are hence defined by the ecological context. 
If plant species adopt comparable roles across different communities, we should expect plant strategies to be more similar within species than across species in the study. 
On the contrary, if species role is flexible, we should expect plants to be able to adopt different strategies in each community such that competition is minimised depending on the context. 

# Methods

Elucidating how the mechanisms and drivers of facilitation scales to the community level and its theoretical consequences has been elusive in part because of the large amounts of data required.
Previous studies focus on the final picture of interspecific pollen transfer that is obtained by examining the pollen deposited on flower stigmas. 
However, as we have described, teasing apart competition for pollinators and evaluating whether there is a trade-off between the number of partners and the mutualistic benefit also requires data that describes the structure of the interactions and the pollen flows that occur within it, while accounting for the traits of the species that integrate the community. 
Here, we examine a comprehensive data set that includes information about pollen deposition, transfer, visitation, species abundance, and species traits across diverse communities in the Argentinian Pampas. 
## Competition vs. facilitation

Our first objective is to identify whether, at the community level, the net outcome of animal-mediated pollination is facilitation or competition. 
A first line of evidence can be found in the relationship between heterospecific and conspecific pollen deposited in plant stigmas [@tur_evaluating_2016]. 
This is so, because heterospecific pollen can be seen as the cost that plants must pay for receiving conspecific pollen from shared pollinators. 
A neggative relationship (when shared pollinators increasse heterospecific pollen loads and decreasse conspecific loads) may indicate that competition for pollination is strong. 
This negative relationship could arrise, for example, when other, more dominant species, monopolizes a limited number of pollinators [@ref], when heterospecific pollen deposited in the stigma interfers with deposition of conspecific pollen [@ref], or when conspecific pollen is lost to heterospecific stigmas [@ref]. 
In contrast, a positive relationship (when sharing pollinators increasses heterospecfic pollen loads but also conspecific pollen deposition) is what should be observed if facilitation is the dominant outcome. 
For example when the abbundance of shared pollinators is positively influenced by floral resources and the positive effects ripple trough the mutualistic network [@ref]. 

A second line of evidence could be obtained by examining the difference between the amount conspecific pollen deposited in flower stigmas *with* and *without* animal mediated pollination. 
It has been shown that self-pollination is favorable to plant reproduction when competition for pollination is strong because it provides insurance against a poor pollination service [@ref]. 
If competition for pollination is sufficiently strong, then the amount of conspecific pollen deposited in flower stigmas when animal pollination is prevented would be higher than the amount deposited when floral visits are allowed. 
The opposite pattern, may indicate that plants benefit from animal mediated pollination, as even a small difference between outcross and self pollen may be important to plant fitness [@ref], particularly so for non-self-fertile species or those in which there is temporary separation in the maturation of the sexes (dichogamy). 

### Data collection

In each of the studied communities, we quantified pollen deposition in a subset of plant species. 
This subset comprised between three and nine of the most common entomophilus species that were flowering during the sampling period while ensuring that the chosen species covered a wide range on a specialization-generalization gradient. 
Briefly, in each of the selected plants, we removed all flowers except buds that were expected to go into inflorescence on the next day. 
Approximately a quarter of these buds were bagged to prevent animal pollination. 
Two days after inflorescence, we counted the pollen grains in the flowers' pistils, classified them between conspecific and heterospecific pollen, and calculated the number of pollen grains per stigma in each flower. 
More details about the study sites and data colletion and laboratory protocols can be found in @marrero_agricultural_2016. 


### Data analysis

Similar to @tur_evaluating_2016, to evaluate the relationship between heterospecific and conspecific pollen, we used a generalised linear mixed model (GLMM) with a Poisson error distribution in each of our communities. 
In these models, the number conspecific pollen grains per stigma was the response variable and the number of heterospecific grains the predictor. 
The slope and intercept of this relationship was allowed to differ between plant species, which were included as a random effect. 
Models were fitted using the function `lmer` from the R package `lme4 1.1-19` [@ref]
To compare the conspecific pollen deposited with and without animal mediated pollination, we performed a Mann-Withney U test of the number of conspecific pollen grains per stigma in un-bagged and bagged flowers for each of the studied species in a community. 

## Data collection

We performed sampling in three locations, each located in a distinct region of the Argentinian Pampas. 
The sampling locations are distributed across a gradient of precipitation with a humid temperate climate in the eastward Flooding Pampas and an drier as we move inland and westward towards the Pampean Grasslands.
Across all sites, annual mean temperatures range between 14.8 and 15.8 Celsius. 

In each location, we sampled two restored and two agricultural fragments. 
However, due to the lack of available sites in the Flooding Pampas, it was only possible to sample one restored site.
We visited each fragment in November 2010 and February 2011. 
In each of these visits we gathered information related to abundance, plant-pollinator visitation, pollen transfer, and pollen deposition.

Abundance was estimated using two 50 m randomly located transects in each fragment. In a first sampling we counted all units of floral attraction that were found in a 2 m wide strip. In a second sampling we counted and collected all floral visitors while walking at a pace of 10 m per minute [@marrero_effect_2014; @memmott_structure_1999]. 
We constructed quantitative visitation networks using information about the floral visitors collected during abundance transects [@marrero_effect_2014]. 
In addition, we also constructed qualitative visitation networks during independent two hour observations of floral visits across each fragment.
We estimated pollen transfer by examining the pollen loads present on the collected floral visitors [@marrero_exotic_2017]. 
Where the pollen count on an individual was estimated to be less than 2,000 grains, we identified every grain to the species level whenever possible and to pollen complexes when it was not. 
When the pollen count was above 2,000 grains, we classified approx. 50% of pollen and total pollen counts were extrapolated [@bosch_plant-pollinator_2009]. 
If more than 10 pollen grains from a plant species were identified, we assumed that pollination service between the plant and the pollinator existed. 
Finally, we analysed pollen deposition in a subset of the plant community [@marrero_agricultural_2016]. 
This subset comprised between three and nine of the most common entomophilus species that were flowering during the sampling period while ensuring that the chosen species covered a wide range on a specialization-generalization gradient. 
In the selected plants we removed all flowers except buds that were expected to go into inflorescence on the next day. 
A quarter of these buds were bagged to prevent animal pollination. 
Two days after inflorescence, we analysed the pollen grains in the flowers' pistils and classified them between conspecific and heterospecific pollen. 

More details about the study sites, data collection, and laboratory protocols can be found in Marrero *et al.* (2014, 2016, and 2017).

## Data analysis

We explore the quantity of pollination service by examining two metrics. 
First, we use the absolute amount of conspecific pollen (pollen counts per stigma) found on flowers open to animal pollination. 
Second, we use the relative amount of conspecific pollen—the difference between the absolute amount and the that from flowers that were bagged. 
We use both metrics simultaneously because, in the one hand, the relative amount acknowledges that autogamus self-pollination can occur and hence might be a better indication of the pollen that has been actually deposited by mutualistic partners. 
On the other, self-pollination is an adaptive trait [@kalisz_benefits_2003] and some plants are not self-fertile, even if their pollen gets deposited on the stigma. 
<!-- Nevertheless, this difference still provides valuable indication of what the animal contribution to reproductive output.  -->
Both metrics complement each other and provide a complete picture of the quantitative aspect of pollination service. 
Similarly, to explore the quality of pollination we examine the relationship that exists between both the absolute and relative amount of conspecific pollen and heterospecific pollen. 
We use the relative amount of heterospecific pollen for our calculations. 
However, because the amount of heterospecific pollen in bagged flowers was negligible, results are almost identical if we use the absolute amounts. 

To study the amounts of deposited pollen, which are needed to construct our metrics of pollination quantity and quality, we use three sets of linear fixed models [using the R package `nlme 3.1-131`, @pinheiro_nlme_2018]—two for conspecific pollen (absolute and relative) and one for heterospecific pollen. 
This modelling approach allowed us to simultaneously explore whether there is a relationship between the benefit of animal pollination and several ecological variables that can affect the mutualistic trade-off. 
In addition, to estimate and incorporate the uncertainty associated to our results we used a combination of multi-model inference and bootstrap resampling. 

Each model set was based on the same one hundred bootstrap resamples of the deposition data. 
We used bootstrap resampling because, as open and bagged flowers were not paired, this allowed us to directly model the *gain* in pollen density rather than density itself. 
In each data replicate we randomly sampled with replacement a number of open flowers within each community-species combination and an equal number of closed flowers. 
We then log-transformed the pollen densities to improve normality and minimise the impact of outliers (Figure \@ref(S-fig:fig-deposition-gain-log-transformed-or-not)), this generally offered better results that a generalised linear model with poisson (or quasiposson) error structure. 
Finally, for the models of the relative amounts of pollen we calculated the difference between the pollen density among treatments. 

As predictors in all model sets, we included species degree, the plant relative abundance, the plant species' share on the pollen pool, and the species' functional originality. 
Species degree was calculated from combining the quantitative and qualitative visitation networks and corresponds to the number of animal species that were observed to visit the plant species in the community where it was sampled. 
The plant's relative abundance we simply aggregated the flower counts for each plant species in the community.
As with the deposition data, we found that log-transforming the abundance counts drastically improved normality (Figure \@ref(S-fig:fig-abundance-log-transformed-or-not)). 
The species' share of the pollen pool was calculated as the proportion of pollen of that species in the pollen counts being carried by pollinators from the community. 

Functional originality is defined as the distance of a species from the community trait average--the centroid of functional space of the community [@laliberte_distance-based_2010; @coux_linking_2016-1]. 
Specifically, we recorded morphological traits that relate to plant type (herb, shrub, climber), life cycle (annual, perennial), flower colouration, and whether the species is native in the study region or not. 
As phenology has been shown to be a key mechanisms of niche differentiation among plants, we also included the plant abundance in each of the survey months (November to March) as an additional trait in our analysis. 
This approach allowed us to integrate the originality in phenological niche alongside the originality on other functional traits. 
All traits were scaled to have a mean of zero and a standard deviation of one prior to calculating the centroid of the functional space. 
The species-specific functional coordinates were calculated using the R package `FD 1.0-12` [@laliberte_distance-based_2010; @laliberte_fd_2014]. 
To facilitate comparison across model estimates, we scaled all independent variables to have a mean of zero and a standard deviation of one.

In all three sets of models we first evaluated a set of candidate random effects which included random intercepts for plant species as well as random slopes for the relationship between degree and pollen gain. 
In addition we also tested for random intercepts that assumed that the plant species grouping was hierarchically nested in either their community, the land use, and the locality. 
We selected the best random structure by comparing the median Akaike Information Criterion (AIC) of evaluated models. 
After determining the best random structure, we then focused on the model's fixed structure composed by our explanatory variables. 
We used multi-model inference to calculate the relative importance of each variable and its mean effect on pollen density.
First we fitted all possible combinations of explanatory variables to the same bootstrap resamples used previously. 
This yielded one hundred models for each "candidate formula"---a particular combination of explanatory variables. 
We then determined the likelihood of each candidate formula and its relative weight by calculating the median $\Delta$AIC of the one hundred models that constitute it. 
As AIC is only useful to compare models with the same response variable, the $\Delta$AIC of each candidate formula was calculated independently for each bootstrap resample.
Calculating the median value of each candidate formula allowed us to estimate the relative importance of the explanatory variables  using the standard methods of adding the median weight of the candidate models that included the variable. 

To calculate the effect that each variable had on the gain of conspecific and heterospecific pollen we sampled from the estimated coefficients in each candidate model proportional to the median likelihood of each set. 
In other words, the estimated coefficients from more likely candidate constitute a larger share of the distribution of coefficients. 
When a variable was not included in a model set we setted the estimate of its coefficient in the model to zero. 
We expect some of our predictors to be correlated to each other. Estimating the models using multiple combinations of explanatory variables also allowed us to evaluate the potential collinearity between them. 
To do so we inspect how the effect of the explanatory variables changed across candidate formulae, when one or more of them were removed in turns. 

After calculating the effects of the ecological variables in heterospecific and conspecific pollen densities, we can now evaluate their effect in the quality and quantity of pollination. 
For the quantity of pollination, we simply calculated the mean value of the distribution of the effect for both the absolute and relative conspecific pollen densities. 
Again, we used bootstrap resampling from the coefficient distributions to calculate the confidence intervals of the mean. 
For the quality of pollination, we calculated the mean difference between a sample of the distribution of the effects of conspecific pollen and a sample of those of heterospecific pollen gain. Similar to the mean effects on the quantity of pollination, we used bootstrap resampling to calculate confidence intervals of the means.estimated by the relationship that exists between the gain in conspecific and heterospecific pollen.

# Results

```{r differences-in-self-pollination}
drake::loadd(significant_gain_global, significant_gain_site)

sgg <- glance_table(significant_gain_global)
prop_sig_global <- sum(sgg$p.value < 0.05) /
  length(significant_gain_global)
sgs <- glance_table(significant_gain_site)
prop_sig_site <- sum(sgs$p.value < 0.05) /
  length(significant_gain_site)

consistency <- dplyr::inner_join(sgg, sgs, by = 'recipient') %>% 
  dplyr::mutate_at(.vars = c('p.value.x', 'p.value.y'), 
            .funs = function(x) x < 0.05) %>%
  dplyr::mutate(consistent = p.value.x == p.value.y) %>%
  dplyr::group_by(recipient) %>%
  dplyr::summarise(p.value.x = mean(p.value.x), 
            consistent = all(consistent),
            n_sites = dplyr::n_distinct(site_name)) %>% 
  dplyr::filter(n_sites > 1)
prop_consistency <- sum(consistency$consistent)/
  nrow(consistency)
```

We found that, for both the model sets of conspecific and heterospecific pollen, the random structure that was best supported by the data was the one that included plant species nested in community as the grouping factors (Table \@ref(S-tab:table-aic-random-models)). 
When looking at the relative amount of conspecific pollen gained, we observe that only `r round(prop_sig_global*100)`% of species in their communities observed a significant increase in conspecific pollen density when comparing open and bagged flowers (Table \@ref(S-tab:table-species-gain-global)).
We also observed a positive relationship between the gain in heterospecific pollen and both the relative and absolute gain of conspecific pollen (Figure \@ref(fig:fig-trade-off-predictions)). 
The slope of the relationship is such that across plant species in their communities heterospecific pollen is gained slower than the absolute amount of conspecific pollen but faster than the amount of conspecific pollen relative to the potential selfing. 

```{r fig-trade-off-predictions, fig.height= 3, fig.width = 6.5, fig.cap = "The interplay between the gain on heterospecific and conspecific pollen. Each dot indicates the predicted median gains for a species in its community (which correspons to the selected random effects in all of our models). The horizontal and vertical bars under each dot indicate the 0.05 bootstrap confidence intervals of the estimates. The solid lines indicate the overall relationship in each of the bootstrap replicates and were calculated by sampling the predictions from 100 conspecific/heterospecific model pairs and estimated the relationship by fitting an standardised major axis (SMA) to each pair (we used the R package smatr 3.4-3, Warton et al. 2012). The dashed line indicates the isoline at which heterospecific pollen is gained at the same rate than conspecific pollen."}
# pdf(width = 6.5, height = 3.25)
  cowplot::plot_grid(
    drake::readd(fig_con_con),
    drake::readd(fig_con_hetero_gain))
  # dev.off()
```

In addition, AIC scores of the candidate model sets indicated that the species' share in the pollen pool, the functional originality, and the species relative abundance were the most important variables determining the gain in both conspecific and heterospecific pollen (Figure \@ref(fig:fig-model-results-global)). 
Note that relative abundance plays a more important role determining the gain in heterospecific pollen than the gain in both the absolute and relative amounts of conspecific pollen.
Surprisingly, the species degree was comparatively less important for predicting pollen gain as it was only included in models with relatively large AICc values (Table \@ref(S-tab:table-aic-fixed-models)).

```{r fig-model-results-global, fig.height = 2, fig.cap = "Distribution of effect estimates of models of conspecific (absolute and relative to the potential selfing) and heterospecific pollen density gain. Results are only shown for models that included explanatory variable. The median relative variable impotance of the four variables for each model set is shown next to each distribution. Variable importance was calculated as the sum of the median Akaike weights (Table \\ref{S-tab:table-aic-fixed-models}) of each candidate model set that includes the variable."}
drake::readd(fig_coefficient_averages)
```

Overall we observed that species degree and abundance had a negative relationship with the gain of both absolute and relative conspecific pollen (Figure \@ref(fig:fig-model-results-global)). 
Functional originality had a weak negative relationship with the absolute gain of conspecific pollen but a small positive effect on the relative gain. 
The plant's share in the community pollen pool, had opposite effects, with a  positive relationship with the absolute gain and a negative relationship with the relative gain of conspecific pollen.
Although relative abundance, degree, and the share in the pollen pool were positively correlated (Figure \@ref(S-fig:fig-correlation)), the effect that each of them had on conspecific pollen was relatively similar among models that included all or just some of these three explanatory variables (Figure \@ref(S-fig:fig-model-results)). 

The gain in heterospecific pollen was negatively related to the species abundance but positively to the species' share in the pollen pool, functional originality, and degree (Figure \@ref(fig:fig-model-results-global)). 
Compared to the models of conspecific pollen, the collinearity in the explanatory variables had a larger impact in the models of heterospecific pollen. 
For instance, when the species abundance was not accounted for, both the species' share in the pollen pool and its functional originality have a negative relationship with the gain in heterospecific pollen (Figure \@ref(S-fig:fig-model-results)). 

<!-- Over all study sites, we found that `r round(prop_sig_global*100)`% of species observed a significant increasse (at the $\alpha = 0.05$ level) in conspecific pollen density when comparing open and bagged flowers (Table \@ref(S-tab:table-species-gain-global)).  -->
<!-- When looking at the differences within sites, we found that only `r round(prop_consistency*100)`% of the species that were present in more than one site had the same response across community (Table \@ref(S-tab:table-species-gain-site)).  -->
Using the results from the models of conspecific and heterospecific pollen gain we then examined the relationship between the explanatory variables and the quantity and quality of pollination (Figure \@ref(fig:fig-average-qual-quant)). 
Degree, had a negative, albeit relatively small, effect on both the quantity and quality of pollination. 
The species relative abundance has a large positive effect to the quality of pollination accompanied by a relatively smaller negative impact on the quantity. 
Finally, the plant's functional originality and its share in the pollen pool have a moderate positive impact on its quantity but also a negative impact on the quality. 

```{r fig-average-qual-quant, fig.height=2, fig.width=3.25, fig.cap="Mean relative effect of ecological variables on the quantity and quality of pollination. Results are shown for values calculated with both the absolute gain in pollen density (abs.) and the gain relative to selfing (rel.). Bootstrap confidence intervals of the mean (at the 0.05 level) are shown in square brackets. "}
drake::readd(fig_average_qual_quant)
```

# Discussion 

* We did find a trade-off between the number of partners a species interacts with and the mutualistic service. Interacting with more species alone is often accompanied by a decreases in both the quantity and quality of pollination. 
* However, the effect and importance of this trade-off is comparatively small when considering other factors that describe the species niche and its ecological settings. 
* We found that animal pollination only brings a small amount of extra conspecific pollen (compared selfing) if any. This relatively small difference is probably very important because, first, some plants simply aren't self-fertile, even if their pollen gets deposited on the stigma. Second, in some flowers there is temporary separation in the maturation of the sexes (dichogamy) that favours cross-pollination as, evolutionary, it has been shown to be a better strategy than cloning. 
* Overall it seems like density seems to be a dominant force behind pollen transfer. If you have many individuals of your species, you get fewer visits (dilution of the pollinators flying around), which is reflected in a smaller amount of deposited pollen. We show that abundant plants also tend to comprise a greater component of the pollen source pool and so the end result is that get more conspecific pollen per visit than heterospecific pollen. 
* Our result of the comparatively weak effect of the mutualistic trade-off is important because it is fundamentally related to the way we understand and mathematically represent mutualistic systems. 
Some of the most widely used theoretical models of pollination account for the negative effects of mutualism by adjusting the mutualistic benefit of an interaction so that it can be negatively affected by the number of mutualistic partners---the species degree. 
This approach, considerably simplifies the mathematical treatment of competition for pollinators. 
However we show that care must be taken because the negative effects of the mutualistic trade-off in both the quantity and quality of pollination can be offset by density mechanisms. Species with large degrees also tend be abundant and prolific pollen producers. 

# Acknowledgements

`r paste(drake::readd(acknowledgements), collapse = "")`

# References
