---
title: "The pollination trade-off"
subtitle: "Supplementary information"
author: "Fernando Cagua, Hugo Marrero, Jason Tylianakis, Daniel Stouffer"
# bibliography: references.bib
csl: ecology-letters.csl
documentclass: book
nocite: | 
  @Xu2003, @Nakagawa2013
output: 
  bookdown::pdf_document2: 
    keep_tex: yes
    toc: false
header-includes:
  - \usepackage{booktabs}
---

\renewcommand\thefigure{S\arabic{figure}}    
\setcounter{figure}{0}   

\renewcommand\thetable{S\arabic{table}}    
\setcounter{table}{0}   

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, 
                      warning = FALSE)
options(xtable.comment = FALSE, 
        xtable.caption.placement = 'bottom',
        xtable.size = 'small',
        booktabs = TRUE)
```

```{r fig-deposition-logtransformed-or-not, fig.height = 2, fig.cap = "Distribution of stigmatic pollen density plotted in (A) a linear scale, and (B) a logarithmic scale (zero values not shown)."}
drake::loadd(fig_pollen_density)

leg <- cowplot::get_legend(fig_pollen_density[[1]])
cowplot::plot_grid(
  remove_legend(fig_pollen_density[[1]]),
  remove_legend(fig_pollen_density[[2]]), 
  labels = LETTERS[1:2], 
  label_size = 8) %>%
  cowplot::plot_grid(leg, ., rel_heights = c(0.1, 0.9), ncol = 1)
```

```{r fig-deposition-gain-logtransformed-or-not, fig.height = 2, fig.cap = "Distribution of the difference on stigmatic pollen density between open and closed flowers for one of the bootstrap replicates used in the model sets. When (A) using directly the gain in pollen density and (B) when pollen density is log transormed prior to calculating the gain. "}
drake::loadd(fig_pollen_density_diff)

leg <- cowplot::get_legend(fig_pollen_density_diff[[1]])
cowplot::plot_grid(
  remove_legend(fig_pollen_density_diff[[1]]), 
  remove_legend(fig_pollen_density_diff[[2]]), 
  labels = LETTERS[1:2], 
  label_size = 8,
  ncol = 2) %>%
  cowplot::plot_grid(leg, ., rel_heights = c(0.1, 0.9), ncol = 1)
```

```{r fig-abundance-logtransformed-or-not, fig.height = 2, fig.cap = "Distribution of plant abundance as (A) raw counts of floral units across communities, and (B) after applying a data transformation in which the counts have been log transformed and scaled to have a mean of zero and a standard deviation of one."}
drake::loadd(fig_abundance)

cowplot::plot_grid(
  fig_abundance[[2]], 
  fig_abundance[[1]],
  rel_widths = c(0.42, 0.58), labels = LETTERS[1:2], label_size = 8)
```

```{r table-species-gain-global, results = 'asis'}
drake::loadd(significant_gain_global, significant_gain_site, sites)

glance_table(significant_gain_global) %>%
  dplyr::select(recipient, estimate, statistic, p.value) %>%
  dplyr::mutate(recipient = paste0('{\\emph{ ', recipient, '}}')) %>%
  dplyr::rename(`p value` = p.value,
         `difference` = estimate,
         `plant species` = recipient) %>%
  knitr::kable(
    caption = "Results of testing the alternative hypothesis that the conspecific pollen density in open flowers is greater than the density in bagged flowers. Tests were performed at the species level (across communities).", 
    escape = FALSE, 
    booktabs = TRUE, 
    linesep = c("")) %>%
  kableExtra::kable_styling(font_size = 8)
```

```{r table-species-gain-site, results = 'asis'}
drake::loadd(significant_gain_global, significant_gain_site, sites)
table_frame <- glance_table(significant_gain_site) %>%
 dplyr::inner_join(sites, by = "site_name") %>%
  dplyr::group_by(recipient) %>%
  dplyr::mutate(n_sites = dplyr::n_distinct(site_name)) %>%
  dplyr::filter(n_sites > 1) %>%
  dplyr::group_by() %>%
  dplyr::mutate(site_name = paste(locality_name, land_use, fragment, sep = ' - '),
         recipient = paste0('{\\emph{ ', recipient, '}}')) %>%
  dplyr::select(recipient, site_name, estimate, statistic, p.value) %>%
  dplyr::rename(`p value` = p.value,
         `difference` = estimate,
         `plant species` = recipient, 
         `community` = site_name)

seps <-dplyr::if_else(table_frame$`plant species` == 
  dplyr::lag(table_frame$`plant species`), "", "\\addlinespace") %>%
  na.omit()

table_frame %>%
  knitr::kable(
    caption = "Results of testing the alternative hypothesis that the conspecific pollen density in open flowers is greater than the density in bagged flowers. Tests were performed at the community level. Only species present in more than one community are shown.", 
    escape = FALSE, 
    booktabs = TRUE, 
    linesep = seps, 
    digits = 3) %>%
  kableExtra::kable_styling(font_size = 8)
```

```{r table-variation-across-sites, results='asis'}
drake::loadd(consp_self)
consp_self %>%
  anova(test = "Chisq")  %>%
  knitr::kable(
    caption = 'Analysis of variance of conspecific pollen density in bagged flowers (self pollination rate).  Density were modelled using a quasipoisson distribution. The model suggests that self-pollination rates are species dependent but not across.',
    booktabs = TRUE) %>%
  kableExtra::kable_styling(font_size = 8)
```

```{r fig-global-community-correlations, fig.height = 2, fig.cap = "Relationship between the model's independent variables calculated at the community level and at the study-wide level."}
drake::readd(fig_community_global_scatter)
```

```{r eval = FALSE, table-aic-random-models, results= 'asis'}
drake::loadd(glanced_random, random_effects)

summ <- glanced_random %>%
  dplyr::filter(var_trans == "log") %>%
  dplyr::group_by(pollen_category, scale, var_trans, model) %>%
  dplyr::mutate(delta_AIC = AIC - min(AIC)) %>%
  dplyr::group_by(pollen_category, scale, var_trans, random_effect) %>%
  dplyr::summarise(AIC_50 = median(delta_AIC), 
            AIC_05 = quantile(delta_AIC, 0.05),
            AIC_95 = quantile(delta_AIC, 0.95), 
            n = n()) %>% 
  dplyr::group_by() %>%
  dplyr::select(-var_trans) %>%
  dplyr::right_join(expand.grid(pollen_category = unique(.$pollen_category),
                        scale = unique(.$scale),
                        random_effect = unique(.$random_effect))) %>% 
  dplyr::mutate_if(is.numeric, round, digits = 1) %>% 
  dplyr::mutate(delta_AIC = paste0(AIC_50, " [", AIC_05, ", ", AIC_95, "], n = ", n), 
         delta_AIC = dplyr::if_else(grepl("NA", delta_AIC), "---", delta_AIC)) %>%
  dplyr::group_by() %>%
  humanize(random_effects = random_effects) %>%
  dplyr::arrange(pollen_category, random_name, scale) %>%
  dplyr::rename(`pollen category` = pollen_category, 
         `random structure` = random_name) 

contingency_table <- xtabs( ~ `random structure` + `scale` + `pollen category` , data = summ) %>% ftable()

# contingency_table <- summ %>%
  # ftable.formula(`random structure` ~ `scale` + `pollen category`)
  
contingency_table[1:nrow(summ)] <- summ$delta_AIC

xtable::xtableFtable(contingency_table, method = "col.compact", 
             caption = paste('Comparison of the different random structures we considered. The table shows median delta AIC values of', dplyr::n_distinct(glanced_random$model), "bootsrap resamples of the data. The 5th and 95th percentile are shown inside square brackets. Communities are defined by individual fragments but ignore the hierarchical arrangement of sampling sites."), 
         label = 'tab:table-aic-random-models')
```

```{r table-global-vs-community-models, results = 'asis'}
drake::loadd(wilcox_glo_com)

wilcox_glo_com %>%
  dplyr::filter(quality %in% c("rmse", "sigma", "r2c", "o2", "nrmse"),
         var_trans == "log") %>%
  dplyr::select(quality, pollen_category, min_scale, estimate, p.value) %>%
  dplyr::rename(`model metric` = quality,
         `pollen type` = pollen_category, 
         `best model set` = min_scale,
         `shift estimate` = estimate, 
         `p value` = p.value) %>% 
  xtable::xtable(digits = -1, caption = "Results of two sample paired Wilcoxon signed rank test comparing different model quality metrics of the model sets using predictors computed across or within communities. Metrics are the root mean square error (rmse), the residual standard deviation (sigma), the conditional r-squared approximation as proposed by Nakagawa and Schielzeth (2013), the omega-squared value as suggested by Xu (2003), and the normalised root-mean-square error of the model sets constructed using predictors computed at the community or study level.", 
         label = "tab:table-global-vs-community-models")
```{r fig-model-results, fig.height = 3.5, fig.cap = "Distribution of effect estimates for models of conspecific and heterospecific pollen density gain. Only results for the models with the most parsimonous fixed effects. "}
drake::readd(fig_all_model_results)
```

# References

